---
title: Performance of Individual Audit Filters in Predicting Opportunities for Improvement
  in Adult Trauma Patients
author: "Josef Al-Khalili"
output:
  word_document: default
  pdf_document: default
  html_document: default
subtitle: "A single-center registry-based retrospective cohort"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
```

# Abstract

## Background

Trauma is responsible for 9% of global deaths annually and has a
significant impact on individuals and society. Quality Improvement (QI)
programs seek to minimize these adverse outcomes using Audit Filters
(AFs) and interdisciplinary Morbidity and Mortality Conferences (MMC).
AFs are based on criteria for standard care and can function as a
screening tool of eligible patient cases for discussion at MMCs. One of
the goals with MMCs is to determine potential Opportunities for
Improvement (OFI).This study aimed to evaluate the performance of
individual AFs in predicting OFIs and to assess their relevance as a
screening tool for Morbidity and Mortality Conferences (MMC).

## Methods

Register data from 8,309 patient cases were collected for analysis. The
sensitivity, specificity and Area Under the Receiver Operating
Characteristics Curve (AUC) was calculated to assess performance. An
empirical bootstrapping test of AUC values was conducted to compare the
performance of each AF to coincidence (AUC = 0.5) and minimum AUC for
diagnostic tests (AUC = 0.8).

## Results

Four out of the 11 AFs assessed had an AUC of 0.5 (p \> 0.05). The
remaining AFs demonstrated a correlation with OFI, with AUC ranging from
0.51 to 0.63. The AF "\> 60 min until first intervention" exhibited the
highest AUC (0.63, 95% CI [0.61-0.65]).

## Conclusion

No individual AF seemed to perform well in predicting OFI, which suggest
rethinking the current screening method for MMCs.

# Introduction

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global], disproportionately affecting younger individuals in
low- and middle-income countries, primarily due to road traffic injuries
[@vos2020global; @roth2018global; @SweTrau2022Arsrapport]. Trauma also
ranks highest in disability-adjusted life-years (DALYs) [@vos2020global]
and increases the risk of mental disorders and cognitive decline
[@kessler1995posttraumatic; @north1999psychological;
@muscatelli2017prevalence; @newnam2014impacts; @visser2017course],
impacting individuals, communities and society at large
[@newnam2014impacts]. Road traffic injuries alone cost approximately 2%
of gross domestic product (GDP) in high-income countries
[@world2014injuries].

To reduce trauma burden, quality improvement (QI) programs have been
established by the World Health Organization (WHO) and the International
Association for Trauma Surgery and Intensive Care (IATSIC)
[@world2009guidelines]. These programs form the core of current trauma
care systems [@American2022resources] and commonly include
multidisciplinary mortality and morbidity conferences (MMC), preventable
death review panels, and the use of AFs also known as quality indicators
[@world2009guidelines]. Despite QI programs, trauma care remains prone
to errors [@nikouline2021errors; @vioque2014classifying;
@o2013opportunities]. Studies have shown a pooled preventable death rate
of 20% from 1990 to 2014, with more recent studies indicating lower
rates [@kwon2014pooled]. For instance, a recent study in a Swedish Level
I trauma centre found a preventable death rate of 4%
[@ghorbani2018analysis].

The MMC is a central QI technique focusing on anonymity, critical
analysis of specific adverse events, recognizing flawed approaches, and
proposing and implementing improvements [@gregor2016morbidity]. Attended
by healthcare professionals involved in trauma care, MMCs discuss themes
like unexpected mortality, morbidity, and errors
[@orlander2003morbidity], fostering error reduction and encouraging
reporting without negative feedback [@lazzara2022morbidity]. Effective
MMC implementation has been shown to increase morbidity and mortality
reporting [@kong2016analysis].

AFs are criteria based on standardized care to enhance trauma care
quality. Deviation from AFs should raise the attention of health care
professionals with the purpose of improving the quality of care
[@American2022resources]. AFs can involve outcomes like "death" or
adherence to guidelines, such as "placement of two large bore
intravenous lines within 15 minutes of arrival."
[@world2009guidelines].Studies have shown mixed results on the
effectiveness of AFs in reducing trauma-related mortality. While some
early studies reported mortality reduction
[@chadbunchachai2001comparative; @chadbunchachai2003study], a 2009
review by Evans et al. [@evans2009audit] found no substantial evidence
supporting AFs' effectiveness. Stelfox et al. [@stelfox2010quality]
identified over 1,500 AFs but noted a lack of a common, evidence-based
set for evaluating trauma care quality, despite the high perceived
usefulness of AFs [@berg2022perceived].

Opportunities for improvement (OFIs) are decided upon review of
individual patient cases [@American2022resources]. They are typically
associated with failures in initial care [@vioque2014classifying] like
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There is lacking evidence for the use of AFs for
improving trauma care [@evans2009audit; @stelfox2010quality] and the
focus has been on mortality as outcome measure
[@chadbunchachai2001comparative; @chadbunchachai2003study]. Furthermore,
the use of AFs has been associated with high frequencies of false
positives, ranging from 24% to 80% [@sanddal2011analysis;
@roy2017learning], and some AFs do not appear to correlate with OFIs at
all [@lewis2017delay].

The aim of this study is to determine the performance of AFs in
predicting OFIs and furthermore assess their relevance as a screening
tool for MMCs.

```{=html}
<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
```
# Methods

## Study design

We conducted a registry-based study using all trauma patients included
in both the Karolinska University Hospital trauma registry and trauma
care quality database between 2012 and 2022 to compare the performance
of AFs in their ability to predict OFI. This study was approved by
Stockholm Research Ethics Review Board, approval number 2021-02541 and
2021-03531.

<!-- Present key elements of study design -->

## Setting

Karolinska University Hospital, a Trauma level I hospital, treats around
1,500 trauma patients annually. If a patient case results in team
activation or the patient has an Injury Severity Score (ISS) of less
than 9 retrospectively, it is recorded in the Karolinska trauma
registry, which reports to SweTrau. This registry includes data on vital
signs, timings, injuries, interventions, and patient demographics as per
the European Consensus Statement (Utstein template)
[@ringdal2008utstein]

The Karolinska trauma registry also contains a local care quality
database with AFs and OFIs, identified by MMCs through consensus
decisions. OFIs are assessed in multiple stages of scrutiny. Notably,
mortality is directly referred to the MMCs in order to assess OFIs and
deciding whether the death was preventable or potentially preventable, a
classification also falling under the purview of OFIs. 

From 2013 to 2017, efforts focused on
identifying adverse outcomes unrelated to mortality. Initially, trauma cases where individually assessed by a
specialized trauma nurse for potential OFIs. In 2017, this procedure
became standard practice, incorporating preliminary evaluations by
specialized trauma nurses. Cases
flagged by an AF or identified by the nurse as potential
care failures underwent a second review by two specialized nurses. Potential OFIs identified in this review
prompted review in a MMC.

```{=html}
<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
```
## Participants

The study included 8,309 patients treated at Karolinska University
Hospital from 2012 to 2022 who underwent OFI screening. Patients under
15 and those not screened for OFIs were excluded due to differences in
clinical management for minors.

```{=html}
<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
```
```{=html}
<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->
```
```{=html}
<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->
```
## Variables and data sources/measurements

The outcome variable is an OFI, as established by the MMCs and treated
as a dichotomous variable with "Yes - At least one OFI identified" and
"No - No OFI identified". The 11 AFs used at Karolinska University
Hospital served as the exposure variable, as shown in Table 1. Each AF
is labeled "original" or "manually created." Manually created AFs were
calculated using trauma registry data, while original AFs were recorded
in the quality database based on criteria beyond the available registry
data.

The data spanning 2012 to 2022 was retrieved from the Karolinska
University Hospital trauma registry and trauma care quality database.
Patient data on vital signs, care processes and interventions, level of
care and time aspects was retrieved from SweTrau while both exposures
(AFs) and outcome (OFI) were retrieved from the Karolinska University
Hospital trauma care quality database. All data was anonymised to
protect patient privacy.

```{=html}
<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
```
## Bias

Consensus decisions at MMC conferences may introduce misclassification
bias. Additionally, AF flagging by specialized nurses is a manual
process, potentially leading to selection bias. Healthcare professionals
attending MMCs may not accurately recall key details impacting OFI
decisions, introducing recall bias. Furthermore, professionals not
present for specific cases might over- or underestimate OFI
significance.

```{=html}
<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
```
## Statistical methods

The study utilized R [@R2022language] for statistical analysis. Patient
characteristics between cases with and without OFI were compared using
the Wilcoxon signed-rank test for continuous variables and the
Chi-squared test for categorical variables.

The predictive performance of AFs was assessed by calculating
sensitivity and specificity for each AF in predicting OFI. Receiver
operating characteristic (ROC) curves were constructed for binary
classification, and the area under the curve (AUC) was computed to
compare AF performance. [@robin2011proc].

Bootstrap resampling (1,000 iterations) was applied to estimate 95%
confidence intervals for sensitivity, specificity, and AUC. The
significance of each AFs' AUC was evaluated against chance (AUC = 0.5)
and the minimum standard for diagnostic tests (AUC = 0.8)
[@nahm2022receiver] using an empirical bootstrap test (1,000
iterations).

Missing values for AFs were treated as negative (i.e., "not flagged").

```{=html}
<!-- 

(a) Describe all statistical methods, including those used to control
for confounding

(b) Describe any methods used to examine subgroups and interactions

(c) Explain how missing data were addressed 

(d) 

Cohort study: If applicable, explain how loss to follow-up was addressed 

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->
```
# Results

##Participants 
After excluding patients under 15 years old and patients not screened for OFIs, a total of 8,309 individuals were included in the
study. 7,797 out of 8,309 (93.8%) patient cases did not have an OFI and
512 out of 8,309 (6.2%) had an OFI. Figure 1 presents a flowchart
showcasing the exclusion and inclusion of the patient cases from the
trauma registry.

Most trauma patients were men (n = 5,755, 69%). Patient characteristics
varied depending on whether patients had an OFI or not. Respectively,
the mean age was significantly higher\
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2)` vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1)`,
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")` and
the ISS was significantly higher
`r inline_text(tableOne, variable = ISS, column = stat_2)` vs
`r inline_text(tableOne, variable = ISS, column = stat_1)`,
`r inline_text(tableOne, variable = ISS, column = "p.value")`. The
number of intubations in the emergency department was significantly
higher
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)`
vs
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1)`,
`r inline_text(tableOne, variable = ed_intubated, level = "No", column = "p.value")`.
Patients with OFI had longer times to definitive treatment from hospital
arrival compared to patients without OFI
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2)`
vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`.
The time to the first CT was also longer in patients with OFIs compared
to patients without
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2)` vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")`.

The number of OFIs were highest in patients admitted to the intensive
care unit
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
and surgical ward
`r inline_text(tableOne, variable = host_care_level, level = "Surgical ward", column = stat_2)`.

Gender `r inline_text(tableOne, variable = Gender, column = "p.value")`,
the mean systolic blood pressure
`r inline_text(tableOne, variable = ed_sbp_value, column = "p.value")`,
death at 30 days
`r inline_text(tableOne, variable = res_survival, column = "p.value")`
and trauma team activation
`r inline_text(tableOne, variable = Tr_Nivå, column = "p.value")` was
not significantly correlated with OFI.

Notably, there was a statistical significant correlation between GCS at
ED and OFI (p \< 0.001).

Table 2 shows the missing data for each AF. As mentioned in the method
section, these values were interpreted as "not flagged" with an AF. The
variables with the highest amount of missing data were "time to
definitive treatment" (n = 5,990, 72.1%) and "time to first CT" (n =
1,012, 12.2%). The smallest amount of missing data was found in
"SBP\<90" (n = 13, 0.2%).

## Performance of Individual Audit Filters

The performance of each AF was determined through a calculation of
sensitivity and specificity which is demonstrated in Table 4.

The number of patient cases flagged by each individual AF varied between
107 for "GCS\<9 and not intubated" and 3907 for "\>30 min until CT". The
AF with the highest sensitivity was "\>30 min until CT" (59.2%, 95% CI
[55.2-63.8]). The lowest sensitivity was seen in "GCS\<9 and not
intubated" (1.2%, 95% CI [0.2-2]). The AF with the highest specificity
was "GCS\<9 and not intubated" (98.7%, 95% CI [98.5-98.9]). The lowest
specificity was seen in "\>30 min until CT" (53.8%, 95% CI [52.7-54.8]).
Notably, the confidence intervals varied between sensitivity and
specificity, with sensitivity displaying a greater interval.

Table 5 summarises the performance of each individual AF in identifying
OFIs in trauma care by AUC. Four of the 11 AFs, "SBP\<90", "dead at 30
days", "GCS\<9 and not intubated" and "CPR and thoracotomy" showed a
performance in predicting OFIs equal to that of coincidence (AUC = 0.5).
The highest AUC was seen in "\>60 min until first intervention" (0.63,
95% CI [0.61-0.65]). All AFs showed significantly lower AUC compared to
the minimum standard for diagnostic tests (AUC = 0.8) (p \< 0.001).

Table 2: Comparison of missing values for each AF. Missing values is
shown as a number and percentage of the total number of patient cases
included in the study (n = 8309).\
`r tableTwo` Table 3: Patient characteristics of individuals with and
without an OFI. Missing values are shown only when they exist.
`r tableOne` Table 4: Comparison of performance for each AF in
predicting OFIs. Performance is defined by sensitivity (%) and
specificity (%). A 95% confidence interval is shown together with each
performance indicator. The number of cases flagged by each filter is
displayed.\
`r tableThree` Table 5: Comparison of performance for each AF in
predicting OFIs. Performance is defined by the area under the receiver
operating characteristics curve. P-values show the significance of AUC
compared to coincidence (AUC = 0.5) and to the minimum standard for
diagnostic tests (AUC = 0.8) [@nahm2022receiver]. The p-values were
calculated using an empirical bootstrap test. Furthermore, a 95%
confidence interval was used. `r tableFour`

```{r fig1, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE, error=FALSE}
knitr::include_graphics("ofi-flowchart.pdf")
```

Figure 1: Flowchart of the patient selection process. Shows the patients
excluded and included in this study.

```{=html}
<!-- 

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->
```
You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this:
`r 2+5`.

# Discussion

The aim of this study was to evaluate the performance of individual AFs
in predicting OFIs and to assess their suitability as a screening tool
for MMCs. This study showed that no AFs was significantly successful in
predicting OFIs. Four AFs ("SBP\<90," "dead at 30 days," "GCS\<9 and not
intubated," and "CPR and thoracotomy") were equivalent to random chance
in their predictive ability. The highest AUC was for "\>60 minutes until
first intervention,". Notably, all AFs performed significantly below the
accepted standard for diagnostic tests (\>0.8).

## Limitations

OFI, while dichotomously defined, includes a diverse set of outcomes.
This convolutes the process of creating AFs that perform well in
predicting OFIs. AFs would favour the identification of some, but not
all, errors.

Each AF was individually related to OFI. Consequently, if an AF
demonstrated a true positive correlation with an OFI, instances where
the same patient case had multiple AFs introduce uncertainty regarding
which specific AF truthfully predicted the OFI.

The number of missing values varied greatly among the AFs. As stated
above, missing values from each AF were coded as "not flagged" by an AF.
This means that some instances where patient cases were not flagged,
might actually meet the AF criteria, but for some reason not documented.
This reduces the validity of some AFs that featured high amounts of
missing data.

##Interpretation Higher ISS, longer times to CT, longer times to first
intervention, and higher levels of care correlated with OFIs, consistent
with a recent study by Albaaj et al. [@albaaj2023patient]. Notably, GCS
at the ED also showed a correlation with OFIs despite identical median
GCS values, likely due to distribution differences outside the IQR.

The AF "GCS\<9 and not intubated" showed no significant ability to
predict OFIs, similarly shown by Willis et al.
[@willis2008interpreting]. Possibly because the AF might be more
reflective of clinical decision rather than quality of care
[@willis2008interpreting]. For instance, An abnormal GCS resulting from
intoxication is managed differently than one caused by a head injury.

Both delay-related AFs: "\>60 until first intervention" and "\>30min
until first CT", showed a correlation with OFI. Comparably, Teixeira et
al. [@teixeira2007preventable] linked delays to preventable death and
Ghorbani et al. [@ghorbani2018analysis] identified treatment and CT
delays as common errors in trauma care. A recent study in Japan by
Yamamoto et. al [@yamamoto2023immediate] found reduced mortality with
whole-body CT scans within 10 minutes of ED arrival.

"Dead at 30 days" and "CPR and thoracotomy" showed no correlation with
OFI. This could be explained by inevitable mortality, thus presenting
with a smaller time frame for mistakes to be made.

AFs generally exhibited a high specificity relative to their sensitivity
which may be attributed to the outcome being rather uncommon, whilst
every individual AF misses the majority of OFIs. This could suggest that
the AFs are too rigid in their criteria. On the contrary, a too
inclusive criteria might flag several patients, without an OFI
[@willis2008interpreting].

## Generalizability

AFs are widely used, but lack consensus on which to use, purpose, and
application [@stelfox2010quality; @stelfox2011evidence;
@berg2022perceived]. They have static criteria and may not adapt to
evolving healthcare challenges, or upon resolution of the quality gaps
they were designed for, reducing their utility over time. This
underscores the need for regular evidence-based reviews of AF
performance [@santana2014development]. However, these reviews might be
cost-prohibitive

While AFs can help identify an OFI, they might overshadow other
contributing factors, leading to over-reliance and potentially failing
to recognize cases outside their scope. [@willis2008interpreting]. For
example, AFs often focus on aspects relevant to doctors, despite the
interdisciplinary nature of trauma care. Lastly, Confounding variables,
such as different treatment times based on injury severity, add to the
uncertainty of AFs' effectiveness. [@willis2008interpreting].

Due to the study being conducted in a single, trauma level I centre in
Stockholm it is unrealistic to assume applicability of these findings at
other hospitals facing unique challenges [@world2009guidelines]. For
instance AFs that benefit low and middle income may not benefit high
income countries and vice versa [@berg2022perceived].

## Future Studies

Given the expanding field of artificial intelligence and machine
learning, new studies should focus on using these tools to screen for
patient cases at risk of an OFI. This will reduce misclassification and
maintain high performance, attributable to an algorithm's capability to
adapt. A recent preprint by Attergrim et al. [@attergrim2023predicting]
found that machine learning models outperformed the AFs currently used
at Karolinska University Hospital in predicting OFIs. Should the use of
AFs remain, they must be routinely updated to reflect current quality
gaps. <!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->

# Conclusion

AFs are widely used as a tool in QI programmes and should reflect the
current gap in trauma care quality. Strong evidence supporting their
efficacy in quality improvement is scarce. In this study, no audit
filter seemed to show promising performance in predicting OFIs with AUC
ranging from 0.5 to 0.63. These results highlight the importance of
continually evaluating the relevance of AFs. However, this process takes
time and resources and using an alternative to AFs is recommended.

# References

Keep? However, according to Sanddal et al. [@sanddal2011analysis] airway
management was associated with OFI, suggesting that an alternative AF
related to airway management might perform differently.

Notably, OFI in this study was not soley a reflection of mortality but
also morbidity. Rather, it could highlight differences in quality gaps
between hospitals. Trauma care is a multidisciplinary, complex and
time-critical health care service provided by hospitals, often trauma
centres. [@American2022resources]. Trauma centres have a significantly
lower mortality rate compared to non-trauma centres, and the maturity of
a trauma centre correlates with lower mortality, demonstrating the
importance of high-quality, specialised trauma care
[@alharbi2021effectiveness; @mann1999systematic;
@candefjord2020mortality]. It has been shown that such review processes
are associated with high-quality trauma care [@santana2014development].
