---
title: Performance of Individual Audit Filters in Predicting Opportunities for Improvement
  in Adult Trauma Patients
author: "Josef Al-Khalili"
output:
  word_document: default
  pdf_document: default
  html_document: default
subtitle: "A single-center registry-based retrospective cohort"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
```

# Abstract

## Background

Trauma audit filters, based on standards of care, serve as screening tools for patient cases to be discussed at interdisciplinary mortality and morbidity conferences. These conferences aim to identify potential opportunities for improvement in the care of trauma patients. This study aimed to assess the effectiveness of individual audit filters in predicting opportunities for improvement and their relevance as screening tools for morbidity and mortality conferences.

## Methods

We conducted a single-centre registry-based retrospective cohort study based on register data from 8,309 patient cases at the Karolinska University Hospital in Sweden. The sensitivity, specificity and Cohen’s kappa coefficient was calculated to assess performance. Bootstrapping of 1,000 iterations was used to calculate 95% confidence intervals.

## Results

Our analysis included 8,309 patients. Two out of the 11 audit filters had a Cohen’s kappa coefficient of 0 and one audit filter had an inverse relation to opportunities for improvement. The remaining audit filters demonstrated a correlation with opportunities for improvement, with Cohen’s kappa coefficients ranging from 0.02 to 0.12. The audit filters “> 60 min until first intervention” and “ISS > 15 and not in ICU” exhibited the highest Cohen’s kappa coefficient (0.12, 95% CI [0.10-0.15]).

## Conclusion

No individual audit filter performed well in predicting opportunities for improvement, which suggest reassessing the current method of screening for mortality and morbidity conferences.

# Introduction

Trauma, defined as a physical injury and the body’s associated response, causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global]and ranks highest in disability-adjusted life-years (DALYs) [@vos2020global],
impacting individuals, communities and society at large
[@newnam2014impacts]. Road traffic injuries alone cost approximately 2%
of gross domestic product (GDP) in high-income countries
[@world2014injuries].

Quality improvement programs form the core of current trauma care systems 
[@American2022resources] and commonly include
 techniques such as multidisciplinary mortality and morbidity conferences and preventable death review panels to improve trauma care quality
[@world2009guidelines]. Despite these programs, trauma care remains prone to errors [@nikouline2021errors; @vioque2014classifying;
@o2013opportunities]. Studies have shown a pooled preventable death rate of 20%, with more recent studies indicating lower rates [@kwon2014pooled]. A recent study in a Swedish Level I trauma centre found a preventable death rate of 4%
[@ghorbani2018analysis].

The mortality and morbidity conferences is a central technique focusing on anonymity, critical analysis of specific adverse events and recognizing flawed approaches followed by proposing and implementing improvements [@gregor2016morbidity]. Attended by healthcare professionals involved in trauma care, mortality and morbidity conferences discuss themes such as unexpected mortality, morbidity and errors
[@orlander2003morbidity], fostering error reduction and encouraging
reporting without negative feedback [@lazzara2022morbidity]. Effective mortality and morbidity conferences implementation has been shown to increase morbidity and mortality reporting [@kong2016analysis].

Audit filters are used to identify cases for mortality and morbidity conferences. These filters are criteria based on standards of care, guidelines or outcomes and function as a screening tool for potential opportunities for improvement. Deviation from audit filters should raise the attention of health care providers with the purpose of improving the quality of care 
[@American2022resources]. Audit filters can involve outcomes such as “death” or adherence to guidelines, such as “placement of two large bore intravenous lines within 15 minutes of arrival.”
[@world2009guidelines].Studies have shown conflicting results on the effectiveness of audit filters in reducing trauma-related mortality. While some early studies reported a mortality reduction following the implementation of audit filters 
[@chadbunchachai2001comparative; @chadbunchachai2003study], a 2009
review by Evans et al. [@evans2009audit]  found no high quality evidence supporting the effectiveness of audit filters. Furthermore, Stelfox et al.  [@stelfox2010quality]
identified over 1,500 audit filters and noted a lack of a common, evidence-based set for evaluating trauma care quality. Despite the uncertainty, the perceived usefulness of audit filters remains high [@berg2022perceived].

Opportunities for improvement are decided upon review of individual patient cases and aim to reflect current trauma care quality gaps  [@American2022resources]. These opportunities are typically associated with failures in initial care [@vioque2014classifying]  like airway management, fluid resuscitation, haemorrhage control and chest injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There is lacking evidence for the use of audit filters for improving trauma care [@evans2009audit; @stelfox2010quality], but the focus has been on mortality as outcome measure 
[@chadbunchachai2001comparative; @chadbunchachai2003study].Moreover, the use of audit filters  has been associated with high frequencies of false positives, ranging from 24% to 80% [@sanddal2011analysis;
@roy2017learning], and some audit filters do not appear to correlate with opportunities for improvement at all [@lewis2017delay].

The aim of this study is to determine the performance of audit filters in predicting opportunities for improvement and assessing their relevance as a screening tool for mortality and morbidity conferences

```{=html}
<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
```
# Methods

## Study design

We conducted a single-centre registry-based retrospective cohort study  of all trauma patients included in both the Karolinska University Hospital trauma registry and trauma care quality database between 2012 and 2022 to compare the performance of audit filters in their ability to predict opportunities for improvement. This study was approved by Swedish Ethical Review Authority, approval number 2021-02541 and 2021-03531.

## Setting

Karolinska University Hospital manages around 1,500 trauma patients annually. The Karolinska Trauma Registry, which reports to the national trauma registry SweTrau, includes all patient cases that result in trauma team activation or who have an Injury Severity Score (ISS) of more than 9. This registry includes data on vital signs, timings, injuries, interventions, and patient demographics as per a European consensus statement
[@ringdal2008utstein]

The trauma centre at the Karolinska University Hospital also keeps a local care quality database with information intended for the mortality and morbidity conferences, including identified triggered audit filters and opportunities for improvement. Opportunities for improvement are assessed in multiple stages of scrutiny. Notably, mortality is directly referred to the mortality and morbidity conferences to assess opportunities for improvement.

From 2013 to 2017, efforts focused on identifying adverse outcomes unrelated to mortality. Initially, trauma cases were individually assessed by a specialized trauma nurse for potential opportunities for improvement. In 2017, this procedure became standard practice, incorporating preliminary evaluations by specialized trauma nurses. Cases flagged by an audit filter or identified by the nurse as potential care failures underwent a second review by two specialized nurses. Potential opportunities for improvement identified in this review prompted review by the mortality and morbidity conference.

```{=html}
<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
```
## Participants

The study included all patients treated at Karolinska University Hospital from 2012 to 2022 who underwent screening for opportunities for improvement. Patients under 15 were excluded due to differences in clinical management.

```{=html}
<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
```
```{=html}
<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->
```
```{=html}
<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->
```
## Variables and data sources/measurements
The outcome was opportunities for improvement, as established by the mortality and morbidity conference, and was treated dichotomously as “Yes - At least one opportunity for improvement identified” and “No - No opportunity for improvement identified”. 

The eleven audit filters used at the Karolinska University Hospital served as the exposure variables, as shown in Table 1. The data spanning 2012 to 2022 was retrieved from the Karolinska Trauma Registry and trauma care quality database. Patient data on vital signs, care processes and interventions, level of care and time aspects was retrieved from SweTrau while both the exposure audit filters and the outcome opportunities for improvement were retrieved from the quality database. 

All data was anonymised to protect patient privacy.

```{=html}
<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->
```
## Bias

Consensus decisions at the mortality and morbidity conferences may introduce misclassification bias. Additionally, audit filters flagging by specialized nurses is a manual process, potentially leading to selection bias. Healthcare professionals attending mortality and morbidity conferences may not accurately recall key details impacting opportunities for improvement decisions, introducing recall bias. Furthermore, professionals, not present for specific cases might over- or underestimate opportunities for improvement significance.
```{=html}
<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
```
## Statistical methods

 for statistical analysis. Patient characteristics for cases with and without opportunities for improvement were compared using the Wilcoxon signed-rank test for continuous variables and the Chi-squared test for categorical variables.
 
The predictive performance of audit filters was assessed by calculating the sensitivity, specificity and Cohen’s kappa for each audit filter in predicting opportunities for improvement.
Bootstrap resampling (1,000 iterations) was applied to estimate 95% confidence intervals for sensitivity, specificity, and Cohen’s kappa coefficient. 

Missing values for audit filters were treated as negative (i.e., “not flagged”).

# Results

##Participants 
After excluding patients under 15 years old and patients not screened for opportunities for improvement, a total of 8,309 individuals were included in the study. 7,797 out of 8,309 (93.8%) patient cases did not have an opportunity for improvement and 512 out of 8,309 (6.2%) had an opportunity for improvement, see Figure 1.

Patients with an opportunity for improvement had significantly higher mean age `r inline_text(tableOne, variable = pt_age_yrs, column = stat_2)` vs `r inline_text(tableOne, variable = pt_age_yrs, column = stat_1)`,
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")`, ISS `r inline_text(tableOne, variable = ISS, column = stat_2)` vs
`r inline_text(tableOne, variable = ISS, column = stat_1)`, `r inline_text(tableOne, variable = ISS, column = "p.value")` and intubations in the emergency department (ED) `r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)`
vs `r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1)`,
`r inline_text(tableOne, variable = ed_intubated, level = "No", column = "p.value")`. Patients with opportunity for improvement had longer times to definitive treatment from hospital arrival `r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2)` vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`.
 and longer times to first CT
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2)` vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1)`,
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")`. The number of opportunities for improvement were highest in patients admitted to the intensive care unit 
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)` compared to general ward `r inline_text(tableOne, variable = host_care_level, level = "General ward", column = stat_2)`. Gender, mean systolic blood pressure, death at 30 days and trauma team activation was not significantly correlated with opportunities for improvement (p>0.05). Notably, there was a statistically significant correlation between GCS at ED and opportunities for improvement (p < 0.001).
The highest amount of missing data was observed for the audit filter "time to
definitive treatment" (n = 5,990, 72.1%). The smallest amount of missing data was found in "SBP\<90" (n = 13, 0.2%).

## Performance of Individual Audit Filters

The number of patient cases flagged by each individual audit filter varied between 107 for “GCS<9 and not intubated” and 3907 for “>30 min until CT”. The highest number of missing values was observed with “>60 min until first intervention” (n = 5990, 72.1%). The lowest number of missing values was observed with “dead at 30 days” (n = 11, 0.1%) and “ISS > 15 and not in ICU” (n = 11, 0.1%). See table 2 for missing values for each audit filter.

The audit filter with the highest sensitivity was “>30 min until CT” (59.2%, 95% CI [55.2-63.8]). The lowest sensitivity was seen in “GCS<9 and not intubated” (1.2%, 95% CI [0.2-2]). The audit filter with the highest specificity was “GCS<9 and not intubated” (98.7%, 95% CI [98.5-98.9]). The lowest specificity was seen in “>30 min until CT” (53.8%, 95% CI [52.7-54.8]). Notably, the confidence intervals varied between sensitivity and specificity, with sensitivity displaying  wider intervals. See Table 4 for the performance of all audit filters using sensitivity and specificity.

Two of the eleven audit filters, “GCS < 9 and not intubated” and “CPR and thoracotomy”, had a Cohen’s kappa coefficient of 0. The highest Cohen’s kappa was seen for the audit filter “>60 min until first intervention” and “ISS > 15 and not in ICU” (0.12, 95% CI [0.10-0.15]). All audit filters showed none to slight correlation with opportunities for improvement. See table 5 for the performance of all audit filters using Cohen’s kappa coefficient. 


```{r fig1, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE, error=FALSE}
knitr::include_graphics("ofi-flowchart.pdf")
```

```{=html}
<!-- 

Participants
------------

(a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

Descriptive data
----------------

(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)

Outcome data
------------

Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures

Main results
------------

(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period

Other analyses 
--------------

Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->
```
You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this:
`r 2+5`.

# Discussion

The aim of this study was to evaluate the performance of individual audit filters in predicting opportunities for improvement and to assess their suitability as a screening tool for mortality and morbidity conferences. This study showed that audit filters had “no” to “slight” ability to predict opportunities for improvement. Two audit filters, “GCS<9 and not intubated” and “CPR and thoracotomy”, showed no predictive ability. The highest Cohen’s kappa coefficient was seen for the audit filter “>60 min until first intervention” and “ISS > 15 and not in ICU”, but it was still low.

## Limitations

Opportunity for improvement, while binarily defined, includes a diverse set of outcomes. This complicates the process of creating audit filters that perform well in predicting opportunities for improvement. Audit filters would favour the identification of some, but not all, errors.

Each audit filter was individually related to opportunities for improvement. Consequently, if an audit filter demonstrated a true positive correlation with an opportunity for improvement, instances where the same patient case had multiple audit filters introduce uncertainty regarding which specific audit filter truthfully predicted the opportunity for improvement.

The number of missing values varied greatly among the audit filters. As stated above, missing values from each audit filter were coded as “not flagged” by an audit filter. This means that instances where patient cases were not flagged might meet the audit filter criteria, but were not documented. This reduces the validity of audit filters that feature high amounts of missing data.

##Interpretation
Higher ISS, longer times to CT, longer times to first intervention, and higher levels of care correlated with opportunities for improvement, consistent with a recent Swedish study by Albaaj et al.[@albaaj2023patient]. Notably, GCS at the ED also showed a correlation with opportunities for improvement despite identical median GCS values, likely due to distribution differences outside the interquartile range (IQR).

The audit filter “GCS<9 and not intubated” showed no significant ability to predict opportunities for improvement, similarly shown by Willis et al. 
[@willis2008interpreting]. Possibly because the audit filter might be more reflective of clinical decisions rather than quality of care 
[@willis2008interpreting]. For instance, an abnormal GCS because of intoxication is managed differently to an abnormal GCS due to head injury.

Both delay-related audit filters: “>60 min until first intervention” and “>30min until first CT”, showed a slight correlation with opportunity for improvement. Comparably, Teixeira et al.[@teixeira2007preventable] linked delays to preventable death andidentified delay to treatment and delays to CT as common errors in trauma care. A recent study in Japan by Yamamoto et. al [@yamamoto2023immediate] observed significantly reduced mortality with whole-body CT scans within 10 minutes of arrival to the ED.

“CPR and thoracotomy” showed no correlation with opportunities for improvement. This could be explained by non-preventable mortality, thus presenting a smaller time frame for mistakes to be made. A similar explanation could be given to the inverse relation between “Dead at 30 days” and opportunities for improvement.

Audit filters generally exhibited a high specificity relative to their sensitivity which may be attributed to the outcome being rather uncommon, whilst every individual audit filter misses the majority of opportunities for improvement. This could suggest that the audit filters are too rigid in their criteria. On the contrary, too inclusive criteria might flag several patients, without opportunities for improvement [@willis2008interpreting].

## Generalizability

Audit filters are widely used, but lack consensus on which to use, purpose, and application [@stelfox2010quality; @stelfox2011evidence;
@berg2022perceived]. They have static criteria and may therefore not adapt to evolving healthcare challenges, or upon resolution of the quality gaps they were designed for, reducing their utility over time. This underscores the need for regular evidence-based reviews of audit filter performance [@santana2014development]. However, these reviews might be cost-prohibitive.

While audit filters can help identify an opportunity for improvement, they might overshadow other contributing factors, leading to over-reliance and potentially failing to recognize cases outside their scope. [@willis2008interpreting].For example, audit filters often focus on aspects relevant to doctors, despite the interdisciplinary nature of trauma care. Lastly, confounding variables, such as different treatment times based on injury severity, add to the uncertainty of audit filters’ effectiveness. [@willis2008interpreting].

Due to the study being conducted in a single trauma centre in Stockholm it is unrealistic to assume applicability of these findings at other hospitals facing unique challenges [@world2009guidelines]. For instance, audit filters that benefit low and middle income may not benefit high income countries and vice versa [@berg2022perceived].

## Future Studies


# Conclusion

Audit filters are widely used as a tool in quality improvement programmes and should reflect the current gap in trauma care quality. In this study, no audit filter showed a significant performance in predicting opportunities for improvement, with Cohen’s kappa coefficient ranging from -0.01 to 0.12. These results highlight the importance of continually evaluating the relevance of audit filters. 

# Acknowledgements
The authors thank all professionals taking part in the monthly mortality and morbidity conferences.

# References
