---
title: Performance of Individual Audit Filters in Predicting Opportunities for Improvement
  in Adult Trauma Patients
author: "Josef Al-Khalili"
output:
  html_document: default
  word_document: default
  pdf_document: default
subtitle: "A single-center registry-based retrospective cohort"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options:
  markdown:
    wrap: 72
---

```{r sourcecode, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE,error=FALSE}
source("main.R")
library(rofi)
library(dplyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
library(boot)
library(knitr)
library(webshot2)


```

# Abstract

## Background

Trauma audit filters, based on standards of care, serve as screening
tools for patient cases to be assessed at interdisciplinary mortality
and morbidity conferences. These conferences aim to identify potential
opportunities for improvement in the care of trauma patients. This study
aimed to determine the effectiveness of individual audit filters in
predicting opportunities for improvement and their relevance as
screening tools for morbidity and mortality conferences.

## Methods

We conducted a single-centre registry-based retrospective cohort study
based on register data collected between 2013 and 2022 from the
Karolinska University Hospital in Sweden. Performance was measured using
sensitivity, specificity and Cohen's kappa coefficient for ten audit
filters.

## Results

Our analysis included 8,309 patients. Audit filters demonstrated
inconsistent agreement between filters and opportunities for improvement
with a Cohen's kappa coefficient ranging from -0.07 (95% confidence
interval [CI], -0.14-0) for "GCS \< 9 and not intubated" and 0.08 (CI,
0.05-0.12) for "Massive transfusion". The Cohen's kappa coefficient for
all audit filters combined was 0.04 (CI, 0.03-0.04).

## Conclusion

No individual audit filter performed well in predicting opportunities
for improvement, which suggests that the current method of screening for
mortality and morbidity conferences should be revisited.

Level III, Prognostic/Epidemiological ofi, audit filter, quality
improvement

# Introduction

Trauma, defined as a physical injury and the body's associated response,
causes 8-9% of global deaths annually [@world2014injuries;
@roth2018global] and ranks highest in disability-adjusted life-years
(DALYs) between ages 10-49 years [@vos2020global], impacting
individuals, communities and society at large [@newnam2014impacts;
@david2022measuring]. Road traffic injuries alone cost approximately 2%
of gross domestic product (GDP) in high-income countries
[@world2014injuries]. Despite advancements in trauma care, preventable
deaths [@kwon2014pooled; @ghorbani2018analysis;
@teixeira2007preventable; @vioque2014classifying] and avoidable medical
errors persist [@nikouline2021errors; @vioque2014classifying;
@o2013opportunities; @teixeira2007preventable]. To address this, trauma
quality improvement programs [@American2022resources], commonly
including mortality and morbidity conferences [@world2009guidelines]
have been implemented. The conferences have shown to foster error
reduction and encourage reporting [@kong2016analysis] without negative
feedback [@lazzara2022morbidity].

Audit filters are criteria used to identify cases for mortality and
morbidity conferences aiming to pinpoint potential opportunities for
improvement [@American2022resources]. Audit filters can involve outcomes
such as "death" or adherence to guidelines, such as "delay to treatment"
[@world2009guidelines]. Opportunities for improvement are typically
associated with failures in initial care [@vioque2014classifying] like
airway management, fluid resuscitation, haemorrhage control and chest
injury management [@sanddal2011analysis; @roy2017learning;
@o2013opportunities]. There is a lack of common, evidence-based set of
audit filters for evaluating trauma care quality [@stelfox2010quality],
and while early studies reported mortality reduction
[@chadbunchachai2001comparative; @chadbunchachai2003study], newer
studies differ [@evans2009audit; @willis2008interpreting]. Moreover, the
use of audit filters has been associated with high frequencies of false
positives, ranging from 24% to 80% in different contexts
[@sanddal2011analysis; @roy2017learning; @ghorbani2018analysis].

Few studies have focused on non-mortality outcome measures, for
evaluating performance of audit filters but the evidence remains
inconsistent [@di2008acscot; @glance2012association] and some audit
filters do not appear to correlate with opportunities for improvement at
all [@lewis2017delay]. However, the perceived usefulness of audit
filters remains high [@berg2022perceived]. The aim of this study is to
determine the performance of individual audit filters in predicting
opportunities for improvement and their relevance as a screening tool
for mortality and morbidity conferences

# Methods

## Study design

We conducted a single-centre registry-based retrospective cohort study
of all trauma patients included in both the Karolinska University
Hospital trauma registry and trauma care quality database between 2013
and 2022 to compare the performance of audit filters in their ability to
predict opportunities for improvement. This study was approved by the
Swedish Ethical Review Authority, approval number 2021-02541 and
2021-03531.

## Setting

Karolinska University Hospital manages around 1,500 trauma patients
annually. The Karolinska Trauma Registry, which reports to the national
trauma registry SweTrau [@SweTrau], includes all patients that result in
trauma team activation or who have an Injury Severity Score (ISS) of
more than 9. This registry includes data on vital signs, timings,
injuries, interventions, and patient demographics as per the European
consensus statement [@ringdal2008utstein]

The trauma centre at the Karolinska University Hospital also keeps a
local care quality database with information intended for the mortality
and morbidity conferences, including triggered audit filters and
identified opportunities for improvement. Opportunities for improvement
are assessed in multiple stages of scrutiny. Notably, patient deaths are
always reviewed to determine preventability and to identify
opportunities for improvement.

During the study period, the screening of patient cases for potential
opportunities for improvement was formalised. Prior to 2017, not all
patient cases were subjected to an initial evaluation by a specialised
nurse. However, following 2017 all patient cases underwent screening.
Importantly, audit filters and an initial evaluation by a specialised
nurse has consistently functioned as the screening tool for mortality
and morbidity conferences to identify potential opportunities for
improvement.

## Participants

The study included all patients treated at Karolinska University
Hospital from 2013 to 2022 and who were screened for opportunities for
improvement. Patients aged under 15 years were excluded.

## Variables and data sources/measurements

The outcome was opportunities for improvement, as established by the
mortality and morbidity conference, and was treated dichotomously as
"Yes - At least one opportunity for improvement identified" and "No - No
opportunity for improvement identified".

The ten audit filters used at the Karolinska University Hospital served
as the exposure variables. For each audit filter, we identified the
cohort of patients in which the filter was applicable, and evaluated
their performance in that cohort: For "ISS \> 15 and no team activation"
and "ISS \> 15 and not in ICU" we evaluated the filters' performance in
patients with an ISS greater than 15; for "GCS \< 9 and not intubated"
we analysed patients with a GCS less than 9; for "\> 30 min until first
intervention" and "\> 60 min until first CT" we selected patients who
underwent a first intervention and those who received a CT respectively.
Lastly, for "No anticoagulants within 72 hours after TBI" we identified
patients with head trauma as defined by the presence of at least one
head injury according to available AIS codes. See Table 1 for

The data was retrieved from the Karolinska Trauma Registry and trauma
care quality database. Patient data on vital signs, care processes and
interventions, level of care and time aspects was retrieved from the
trauma registry whereas the exposure, audit filters and the outcome,
opportunities for improvement were retrieved from the trauma care
quality database.

## Statistical methods

We used R [@R2022language] for statistical analysis. Patient
characteristics for cases with and without opportunities for improvement
were compared using the Wilcoxon signed-rank test for continuous
variables and the Chi-squared test for categorical variables. The
performance of audit filters in predicting opportunities for improvement
was assessed by calculating the sensitivity, specificity and Cohen's
kappa. A Cohen's kappa of 0-0.2 was interpreted as "no" agreement while
a value from 0.2 to 1 was interpreted as "minimal" up to "perfect"
agreement [@mchugh2012interrater]. Bootstrapping with 1,000 iterations
was applied to estimate 95% confidence intervals for all performance
metrics. We selected cohorts relevant to each filter: For "ISS \> 15 and
no team activation" and "ISS \> 15 and not in ICU" we examined patients
with an ISS greater than 15. For "GCS \< 9 and not intubated," we
focused on patients with a GCS less than 9. Lastly, for "No
anticoagulants within 72 hours after TBI" we identified patients with
head trauma as defined by the presence of at least one head injury
according to available AIS codes. Missing values were numbered but
ignored in the analysis for example those without a registered blood
pressure was ignored when analysing "SBP \< 90".

# Results

After excluding patients younger than 15 years and patients not screened
for opportunities for improvement, a total of 8,309 individuals were
included in the study. Out of these 8,309 patients, 512 (6.2%) patients
had an opportunity for improvement, see Figure 1.

Patients with an opportunity for improvement had significantly higher
median age,
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")`
years vs
`r inline_text(tableOne, variable = pt_age_yrs, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = pt_age_yrs, column = "p.value")`)
years, median ISS
`r inline_text(tableOne, variable = ISS, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")`
vs
`r inline_text(tableOne, variable = ISS, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = ISS, column = "p.value")`) and
intubations in the emergency department (ED),
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_2)`
patients vs
`r inline_text(tableOne, variable = ed_intubated, level = "Yes", column = stat_1, pattern = "{N} ({p}%;")`
`r inline_text(tableOne, variable = ed_intubated, column = "p.value")`)
patients. Patients with opportunity for improvement had longer times to
first major intervention from hospital arrival median
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")`
minutes vs
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = dt_ed_emerg_proc, column = "p.value")`)
minutes. and longer times to first CT, median
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_2, pattern = "{median} (IQR {p25}, {p75})")`
minutes vs
`r inline_text(tableOne, variable = dt_ed_first_ct, column = stat_1, pattern = "{median} (IQR {p25}, {p75};")`
`r inline_text(tableOne, variable = dt_ed_first_ct, column = "p.value")`)
minutes.

The frequency of opportunities for improvement was highest in patients
admitted to the intensive care unit,
`r inline_text(tableOne, variable = host_care_level, level = "Intensive care unit", column = stat_2)`
patients, compared to general ward
`r inline_text(tableOne, variable = host_care_level, level = "General ward", column = stat_2)`.
Notably, there was a statistically significant correlation between GCS
at the ED and opportunities for improvement (p \< 0.001). Gender, mean
systolic blood pressure, death at 30 days and trauma team activation
were not significantly correlated with opportunities for improvement
(p\>0.05). The highest amount of missing data was observed with "time to
first major intervention", which was missing for 5,990 patients (72.1%),
partly because the absence of an initial CT or major intervention due to
the clinical management being observed resulted in their registration as
missing. See Table 2 for details of selected patient characteristics.

## Performance of Individual Audit Filters

The number of patient cases flagged by each individual audit filter
varied between (n = 107, 27.4%) for "GCS \< 9 and not intubated" and (n
= 3907, 53.5%) for "\>30 min until CT". The highest number of missing
values was observed with "\> 60 min until first intervention" (n = 5990,
72,1%). The lowest number of missing values was observed with "dead at
30 days" and "ISS \> 15 and not in ICU" (n = 11, 0.1%). The largest
cohort was observed with "Dead at 30 days" (n = 8298, 99.9%) and the
smallest cohort was observed with "GCS \< 9 and not intubated" (n = 390,
4.7%). In total 5977 out of 8309 (71.9%) patient cases were flagged by
at least one audit filter. See Table 1 for the cohort size, number of
missing values and number of flagged patient cases for each audit
filter.

The audit filter with the highest sensitivity was "\> 60 min until first
intervention" (83.3%, 95% CI, 79.2-87.5). The lowest sensitivity was
seen in "SBP \< 90" (8%, 95% CI, 5.6-10.2) and "Dead at 30 days" (8%,
95% CI, 5.6-10.2). The audit filter with the highest specificity was
"Massive transfusion" (95.8%, 95% CI, 95.4-96.3). The lowest specificity
was seen in "ISS \> 15 and no team activation" (21.6%, 95% CI,
19.9-23.2). The sensitivity for all audit filters combined was 92.6%
(95% CI, 90.5-95) and the specificity was 29.4% (95% CI, 28.4-30.5).

The audit filter with the highest Cohen's kappa was "Massive
transfusion" (0.08, 95% CI, 0.05-0.12). The smallest Cohen's kappa was
seen in "GCS \< 9 and not intubated" (-0.07, 95% CI, -0.14-0. The
Cohen´s kappa for all audit filters combined was 0.04 (95% CI,
0.03-0.04). See Table 3 for the performance of all audit filters using
sensitivity, specificity and Cohen's kappa.

```{r fig1, echo=FALSE, fig.align="center", warning=FALSE, message=FALSE, error=FALSE}
knitr::include_graphics("ofi-flowchart.pdf")
```

# Discussion

This study shows that audit filters, individually or in combination,
were not effective in predicting opportunities for improvement. As a
result, the development of a new screening method to better identify
potential opportunities for improvement is advised.

Few studies have examined the link between individual audit filters and
opportunities for improvement. Given the numerous different audit
filters, comparing our findings to studies that utilize a different set
of audit filters proves challenging. The delay-related audit filters in
this study: "\> 60 min until first intervention", "\> 30 min until first
CT" did not predict opportunities for improvement. Interestingly, time
to definitive treatment and time to first CT was associated with
opportunities for improvement in the patient characteristics analysis,
similar to a previous study by Albaaj et al. [@albaaj2023patient] on the
association between patient factors and opportunities for improvement.
However, while statistically significant, the observed effect size
suggests a diminishable clinical relevance. Similarly, one study by
Lewis et al. [@lewis2017delay] found that delays to the operating room
exceeding 2 hours from emergency department arrival did not individually
predict opportunities for improvement.

Furthermore, the audit filter "GCS \< 9 and not intubated" showed no
correlation with opportunities for improvement in our study. Comparably,
Willis et al. [@willis2008interpreting] found no association between
"GCS below 9 and not intubated" and mortality or length of hospital stay
but it was associated with shorter stays in the ICU. Contrary, Glance et
al. [@glance2012association] found that "No definitive airway in a
patient with GCS score ≤ 8" was associated with increased mortality risk
but paradoxically a reduced risk of major complication and death
combined. This was concluded as reflective of differences in clinical
decision rather than quality of care [@willis2008interpreting;
@glance2012association]. For instance, an abnormal GCS because of
intoxication is managed differently to an abnormal GCS due to head
injury. Furthermore, Willis et al. found that "failure to activate
trauma team at major trauma service" was not associated with poor
outcomes, similar to our findings.

Di Bartolomeo et al. [@di2008acscot] studied the association between
four audit filters ("Ambulance scene time \>20 min", "Patient with a
Glasgow coma scale score (GCS) \< 13 who does not receive a CT scan of
the head within 2 h of arrival at the hospital", "Patients with epidural
or subdural haematoma receiving craniotomy \>4 h after arrival at ED,
excluding those performed for intracranial pressure monitoring",
"Abdominal, thoracic, vascular or cranial surgery performed \>24 h after
arrival") and 30-day mortality and 6-month disability. One audit filter:
"abdominal, thoracic, vascular or cranial surgery performed \>24 h after
arrival" showed contradictory outcomes with reduced mortality but
increased disability. This may be explained by aggressive trauma care
producing additional survivors with poor functional outcomes or
underestimating the mortality rate beyond 30 days [@di2008acscot]. The
rest of the audit filters showed no significant association with
outcomes.

A study by Chadbunchachai et al. [@chadbunchachai2001comparative]
reported a reduction in preventable deaths following the implementation
of audit filters, with similar results in a subsequent study
[@chadbunchachai2003study]. In 2009 Evans et al. [@evans2009audit]
conducted a review to evaluate the effectiveness of audit filters in
improving trauma care quality but found no studies of sufficient
methodological quality for inclusion. The studies by Chadbunchachai et
al. were excluded due to insufficient pre- and post-implementation data
points. They concluded that there is no high-quality evidence supporting
the use of audit filters in improving trauma care. Similarly, Stelfox et
al. [@stelfox2010quality] found no clearly defined, evidence-based set
of audit filters. A more recent study by Berg et al. [@berg2024effects]
also observed decreased mortality after the implementation of audit
filters in India. Furthermore, the perceived usefulness of audit filters
in trauma quality improvement among health care professionals remains
high [@berg2022perceived; @santana2014development].

There are some limitations in this study. Firstly, it is unlikely that
the specific associations presented in this study are generalizable to
other hospitals with similar systems [@di2008acscot;
@glance2012association], but the broader finding that audit filters are
poorly associated with opportunities for improvement is likely to be
externally valid [@world2009guidelines]. Secondly, instances where the
same patient case had multiple audit filters flagged introduce
uncertainty regarding which specific audit filter truthfully predicted
an opportunity for improvement, introducing a level of confounding.
Finally, the number of missing values differed significantly among the
audit filters, resulting in varying degrees of validity for each audit
filter.

The static nature of audit filters inherently limit their adaptability
to emerging healthcare challenges and resolved quality gaps, reducing
their utility over time [@glance2012association]. Therefore, regular
evidence-based reviews are needed to ensure performance
[@santana2014development], though such reviews may be cost-prohibitive.
Furthermore, they may overshadow other contributing factors to
opportunities for improvement, i.e confounding variables
[@glance2012association; @willis2008interpreting].

In summary, audit filters are widely used as a tool in quality
improvement programmes and should reflect current gaps in trauma care
quality. In this study, no individual audit filter or their combination
were successful in predicting opportunities for improvement. Given these
results, in combination with the lack of high-quality evidence showing
that implementing audit filters improve patient outcomes, future studies
should focus on developing new tools with higher predictive ability to
identify opportunities for improvement.

# Acknowledgements

The authors thank all professionals taking part in the monthly mortality
and morbidity conferences.

# References
